\documentclass[manuscript, screen, review, nonacm]{jair}

\usepackage{amsmath}
\usepackage{array}
\usepackage{booktabs}
\usepackage{float}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{url}

\settopmatter{printacmref=false, printccs=false}
\makeatletter
\let\@ACM@checkaffil\relax
\makeatother

\setcopyright{none}
\JAIRAE{To be assigned}
\JAIRTrack{}
\acmVolume{0}
\acmArticle{0}
\acmMonth{6}
\acmYear{2026}

\RequirePackage[
  datamodel=acmdatamodel,
  style=acmauthoryear,
  backend=biber,
  giveninits=true,
  uniquename=init
  ]{biblatex}

\renewcommand*{\bibopenbracket}{(}
\renewcommand*{\bibclosebracket}{)}
\addbibresource{references.bib}

\newcommand{\pc}{\mathrm{PC}}
\newcommand{\pe}{\mathrm{PE}}
\newcommand{\imp}{\mathrm{I}}
\newcommand{\reg}{\mathrm{R}}
\newcommand{\nrb}{\mathrm{NRB}}

\title[Item-Level Replacement Audit]{An Item-Level Audit Protocol for Error Persistence and Regression Burden in Small Language Model Replacement}

\author{Dhananjay Singh Chauhan}
\authornote{Corresponding author.}
\email{dschauhan08.me@gmail.com}
\affiliation{%
  \institution{Independent Researcher}
}

\renewcommand{\shortauthors}{Chauhan}

\begin{document}

\begin{abstract}
Replacement decisions for small language models are often justified by aggregate benchmark gains. That practice can hide a deployment-relevant distinction: a candidate model may improve the mean score while newly failing on items that the current model answered correctly. This software-tool article presents a paired item-level replacement-audit protocol. For each shared labeled item, the protocol separates persistent correct answers, corrected errors, regressions, and persistent errors, then reports correction rate, regression mass, churn mass, error persistence, and normalized regression burden.

The contribution is a practical reporting protocol and public artifact, not a scaling law or a causal account of model families. The primary evidence is a parser-independent WILD correctness-only use case with 276,685 item-level records for five Qwen and Llama models up to approximately four billion parameters across ten tasks. In this public setting, Qwen2.5-0.5B to Qwen2.5-3B improves aggregate correctness by 22.2 percentage points while changing 40.9\% of item outcomes; Llama 3.2 1B to 3B improves by 17.8 points while changing 35.3\% of outcomes; and the near-parity Qwen2.5-1.5B to Qwen2.5-3B comparison improves by 5.7 points while changing 30.4\% of outcomes. A secondary archived-output MMLU-Pro use case reconstructs 3,008 source items across 14 model-output runs and demonstrates raw-response audit tooling, but remains diagnostic because complete inference manifests and human parser validation are not yet available. The core claim is narrow: benchmark-based replacement decisions should report item-level corrections, regressions, persistent errors, and churn alongside aggregate accuracy.
\end{abstract}

\maketitle
\begin{center}
\textbf{Article type: Software Tool Article / Methods and Software Artifact}
\end{center}
\begin{center}
\begin{minipage}{0.92\linewidth}
\small
\textbf{AI disclosure statement.} AI-assisted editorial and software-development tools were used during manuscript drafting, code review, and artifact preparation. The author reviewed, edited, verified, and takes responsibility for the final manuscript, code, analyses, and claims.
\end{minipage}
\end{center}
\thispagestyle{plain}
\pagestyle{plain}

\section{Introduction}

Small language models are attractive deployment targets because they can satisfy memory, latency, privacy, and cost constraints that make larger models impractical. In that setting, replacement is an ordinary engineering decision. A team has a current model, evaluates a candidate model, and must decide whether the candidate is a better component for the same task. Benchmark accuracy is often the first summary used for this decision.

Aggregate accuracy is a useful summary, but it is not a replacement audit. The same net gain can arise from very different item-level behavior. A candidate may correct many current-model errors while introducing few new failures, or it may achieve a similar mean score after replacing one set of correct items with another. The second case matters in applications where previously reliable cases carry operational risk, such as routing documents, flagging compliance cases, or triaging support requests. A higher average score does not tell a user which old errors remain, which old errors were fixed, or which old successes were lost.

This paper treats model replacement as a paired item-level accounting problem. Given two models evaluated on the same labeled item identifiers, each item falls into one of four cells: persistent correct, corrected error, regression, or persistent error. From these counts, the audit reports aggregate gain together with improvement mass, regression mass, churn mass, error persistence, correction rate, and normalized regression burden. The result is not a new training method or a mechanistic explanation of model behavior. It is a reporting protocol for a practical question: what changes, item by item, when one small model is used in place of another?

The empirical design follows that modest claim. The primary demonstration uses WILD, a public item-level correctness matrix, because binary correctness records avoid the answer-extraction problem that arises with raw language-model completions. The retained WILD subset contains five Qwen and Llama models up to approximately four billion parameters across ten tasks. It is not presented as a representative census of all small models. It is used to show, in a public parser-independent setting, that aggregate gains and item-level churn are distinct quantities.

The paper also includes an archived-output MMLU-Pro case study. That case study is useful because it shows the additional work required when the audit begins from raw model responses: source-row reconstruction, answer extraction, scoring-mode sensitivity, parser-audit sampling, and run-provenance checks. It is not treated as confirmatory evidence for parser-dependent claims, because the original run logs do not contain complete inference manifests and the manual parser audit is not complete. This evidence split is intentional. WILD carries the core item-level correctness claim; MMLU-Pro demonstrates the raw-output workflow and its remaining validation requirements.

The central question is therefore deliberately narrow:

\begin{quote}
When a candidate small language model improves a benchmark score, how much of the shared item-level correctness pattern is corrected, regressed, or left unchanged?
\end{quote}

The paper makes four contributions:

\begin{enumerate}
  \item A reusable replacement-audit protocol for paired item-level correctness, including correction rate, regression mass, churn mass, error persistence, and normalized regression burden.
  \item A clear distinction between deployment-path audits and standardized all-pairs diagnostics, so directional metrics are interpreted only with respect to a specified current model and candidate model.
  \item A public WILD correctness-only demonstration showing that aggregate improvements can coincide with substantial item-level churn without relying on private completions or answer parsing.
  \item A diagnostic archived-output MMLU-Pro case study with source-item hashes, scoring-mode sensitivity checks, parser-audit tooling, and explicit limits on answer extraction and run provenance.
\end{enumerate}

This submission is framed as a software-tool article: the main object is the audit protocol and public artifact, with WILD serving as the primary public use case and MMLU-Pro serving as a diagnostic raw-output workflow.

The paper does not claim that prediction churn is new, that one model family has a general property, or that the reported values define a scaling law. Its claim is operational: when benchmark scores are used to justify replacing a model, the report should include corrected errors, regressions, persistent errors, and churn, not aggregate accuracy alone.

\section{Related Work}

\paragraph{Beyond aggregate accuracy.}
NLP evaluation has repeatedly shown that aggregate scores can hide important behavior. CheckList introduced behavioral tests to expose failures hidden by held-out accuracy \parencite{ribeiro2020checklist}. HELM argues for transparent, multi-metric model evaluation rather than single-number comparisons \parencite{liang2022helm}. The present audit follows the same evaluation philosophy, but focuses on paired item-level changes between a current model and a candidate replacement model.

\paragraph{Prediction churn and regression bugs.}
Prediction churn studies how model predictions change across retraining or deployment iterations \parencite{fard2016churn}. Regression-bug work in NLP model updates measures negative flips: cases where a newer model newly fails examples an older model handled correctly \parencite{xie2021regression}. Unlike work that proposes training or update methods to reduce regressions, this paper proposes a reporting and audit artifact for replacement decisions. The present work does not claim to invent churn or regression measurement. Its narrower contribution is to adapt this accounting to small language model replacement and to keep beneficial churn, harmful churn, and persistent old errors separate in the artifact and reporting.

\paragraph{Regression testing and deployment audit.}
The audit is close in spirit to software regression testing. Before replacing a component, users need to know which previously passing cases now fail. In model evaluation, the analogous question is which items were previously correct and are now wrong. This is a deployment-facing question rather than a claim about training mechanisms.

\paragraph{Production readiness and documentation.}
Production ML work emphasizes systematic testing, monitoring, and documentation rather than isolated aggregate scores \parencite{sculley2015technicaldebt,breck2017mltestscore}. Dataset and model documentation frameworks likewise stress provenance, intended use, and evaluation context \parencite{gebru2021datasheets,mitchell2019modelcards}. The present artifact follows that orientation by separating source provenance, evidence tier, scoring status, and claim use.

\paragraph{Stability, forgetting, and underspecification.}
The replacement problem also connects to work on catastrophic forgetting \parencite{kirkpatrick2017overcoming} and underspecification \parencite{damour2022underspecification}. Those literatures study different mechanisms and failure modes. Here, the object is narrower: observable item-level correctness changes between archived model outputs or public correctness records.

\paragraph{Benchmarks and public item-level resources.}
Benchmark suites such as MMLU, MMLU-Pro, and BIG-bench provide broad task-level measurement \parencite{hendrycks2021mmlu,wang2024mmlupro,srivastava2022bigbench}. WILD provides public item-level correctness records that can be used for correctness-only replacement audits \parencite{krumdick2026wild}. Aggregate leaderboards and model cards are useful for context, but they cannot identify which old errors were corrected or which previously correct items regressed.

\section{Replacement-Audit Protocol}

\subsection{Inputs and evidence tiers}

The audit requires paired item-level outcomes for two models on the same labeled item identifiers. The strongest input is raw model output plus a validated scoring procedure, meaning either benchmark-provided predictions or a completed human-labeled parser audit. A weaker but still useful input is a public binary correctness matrix. Aggregate leaderboards alone are insufficient.

\begin{table}[t]
\centering
\small
\begin{tabular}{p{0.12\linewidth}p{0.35\linewidth}p{0.39\linewidth}}
\toprule
Tier & Contents & Permitted use \\
\midrule
A & Raw prompts, raw completions, labels, item IDs, and validated scoring & Full audit, parser analysis, format-failure analysis \\
B & Item IDs, model IDs, and binary correctness & Correctness-only churn, regression, persistence, and correction metrics \\
C & Aggregate benchmark scores only & Context only; no item-level replacement claims \\
D & Gated, unavailable, incompatible, or unverified data & Excluded until access and terms are clear \\
\bottomrule
\end{tabular}
\caption{Evidence tiers for replacement accounting. Only Tier A and Tier B sources support item-level churn metrics.}
\label{tab:evidence-tiers}
\end{table}

\begin{table}[t]
\centering
\small
\begin{tabular}{p{0.30\linewidth}p{0.18\linewidth}p{0.42\linewidth}}
\toprule
Evidence component & Status & Claim use \\
\midrule
WILD item-level correctness & Supported & Core replacement-accounting evidence for persistence, regression mass, and churn over paired correctness records. \\
MMLU-Pro source reconstruction & Supported & Provenance evidence that all 3,008 evaluated item identifiers map to the stated MMLU-Pro snapshot. \\
MMLU-Pro raw-output accounting & Diagnostic only & Demonstrates the raw-output workflow and artifact-risk checks, but remains parser-dependent. \\
Parser-validation claims & Not claimed & Human labels are incomplete, so the paper does not report parser agreement as validation. \\
Aggregate benchmark context & Context only & Positions model capability; does not support item-level replacement or churn claims. \\
\bottomrule
\end{tabular}
\caption{Claim ledger for the current GitHub/SSRN release. Parser-dependent MMLU-Pro claims are explicitly excluded from the core claim set until human validation is complete.}
\label{tab:claim-ledger}
\end{table}

\subsection{Four-cell decomposition}

For a directed comparison from current model $A$ to candidate model $B$ over the same $N$ items, each item belongs to exactly one cell:

\begin{itemize}
  \item persistent correct ($\pc$): both models are correct;
  \item improvement or corrected error ($\imp$): $A$ is wrong and $B$ is correct;
  \item regression ($\reg$): $A$ is correct and $B$ is wrong;
  \item persistent error ($\pe$): both models are wrong.
\end{itemize}

The aggregate accuracy delta is
\begin{equation}
\Delta = \frac{|\imp|-|\reg|}{N}.
\end{equation}

The audit metrics are
\begin{align}
\text{Improvement mass} &= \frac{|\imp|}{N}, \\
\text{Regression mass} &= \frac{|\reg|}{N}, \\
\text{Churn mass} &= \frac{|\imp|+|\reg|}{N}, \\
\text{Error persistence} &= \frac{|\pe|}{|\pe|+|\imp|}, \\
\text{Correction rate} &= \frac{|\imp|}{|\pe|+|\imp|}.
\end{align}

Let the current model's error rate be
\begin{equation}
 e_0 = \frac{|\pe|+|\imp|}{N}.
\end{equation}

The normalized regression burden is
\begin{equation}
 \nrb = \frac{|\reg|/N}{e_0}.
\end{equation}

If $|\pe|+|\imp|=0$, the current model has no observed errors on the shared item set. Error persistence, correction rate, and $\nrb$ are then undefined and are reported as \texttt{NA}, not as zero. Category summaries with zero current-model errors remain included for mass metrics such as improvement mass, regression mass, and churn mass, but old-error-normalized ratios are reported as \texttt{NA}.

For an accuracy-improving directed comparison, $\nrb$ is a practical normalization of regression mass by the current model's observed error rate. It can also be read descriptively as the gap between the no-regression upper bound on old-error persistence and observed old-error persistence. This interpretation is descriptive; it is not a theoretical novelty claim and does not identify a causal mechanism.

\subsection{Directed paths versus all-pairs diagnostics}

The metrics are directional. A comparison $A \rightarrow B$ answers a different deployment question than $B \rightarrow A$. A deployment audit should therefore use the actual proposed replacement path whenever it is known.

When only a set of model outputs is available, a standardized all-pairs diagnostic can still be useful. In that case, the ordering rule must be explicit and should not be mistaken for vendor release chronology. In the MMLU-Pro case study below, directed pairs are ordered by parameter count, then generation string, then model label. Comparisons marked as chronology unclear are treated as diagnostic pairs, not as true deployment histories.

\subsection{Minimal audit-card algorithm}

\begin{table}[t]
\centering
\small
\begin{tabular}{p{0.38\linewidth}p{0.52\linewidth}}
\toprule
Field & Purpose \\
\midrule
Current and candidate model accuracy & Aggregate movement \\
Accuracy delta & Net gain or loss \\
Improvement mass & Share of all evaluated items that the current model missed and the candidate model corrected \\
Regression mass & Share of all evaluated items that the current model answered correctly and the candidate model newly misses \\
Churn mass & Total share of changed item outcomes \\
Error persistence and correction rate & How much of the current error set remains \\
Normalized regression burden & Practical normalization of regression mass relative to current-model errors \\
Top regressing categories or clusters & Where manual review should start \\
Path note & Whether the directed pair is a real replacement path or a diagnostic comparison \\
\bottomrule
\end{tabular}
\caption{Minimal replacement-audit card emitted by the artifact.}
\label{tab:audit-card}
\end{table}

\begin{enumerate}
  \item Align current-model and candidate-model records by item identifier.
  \item Convert each model's output to a binary correctness label under a documented scoring procedure.
  \item Count $\pc$, $\imp$, $\reg$, and $\pe$.
  \item Compute $\Delta$, improvement mass, regression mass, churn mass, error persistence, correction rate, and $\nrb$.
  \item Emit an audit card with optional review flags for high churn, high regression mass, high $\nrb$, near-parity churn, and negative-delta swaps.
\end{enumerate}

\subsection{Statistical posture}

The audit is primarily descriptive. McNemar-style paired tests are appropriate when comparing two classifiers on the same items \parencite{mcnemar1947note,dietterich1998approximate}, but replacement accounting should not treat shared-endpoint model pairs as independent evidence. The artifact therefore reports mass and churn metrics as the primary outputs and uses inferential checks only as clearly labeled diagnostics.

\section{Implementation and Operation}

The public artifact is structured as a small analysis package rather than a monolithic notebook. Each command writes CSV, JSON, and paper-table outputs that can be inspected independently. Table~\ref{tab:module-map} summarizes the main modules, their input requirements, and their role in the paper.

\begin{table}[t]
\centering
\small
\begin{tabular}{p{0.24\linewidth}p{0.25\linewidth}p{0.27\linewidth}p{0.14\linewidth}}
\toprule
Module & Input & Output & Paper use \\
\midrule
WILD ingest & WILD item-level correctness records & Normalized records, pairwise replacement metrics, task-dispersion tables & Core \\
Evidence registry & Source registry entries & Evidence-tier map and claim-use table & Claim boundary \\
MMLU-Pro manifest & Derived MMLU-Pro item files and source parquet & Redacted source-row hashes and manifest status & Provenance \\
Paper metrics & Derived MMLU-Pro correctness summaries & Pairwise metrics, audit cards, appendix tables & Diagnostic \\
Parser-impact audit & Private parser-audit labels, when available & Parser-impact report and claim gate & Diagnostic gate \\
Release hygiene & Public release directory & Leakage scan for raw responses, benchmark text, private samples, and local paths & Release safety \\
\bottomrule
\end{tabular}
\caption{Software modules used by the public artifact. The WILD module validates the pairwise accounting layer; raw-output parser validation is a separate diagnostic layer and is not claimed complete.}
\label{tab:module-map}
\end{table}

The public package targets Python 3.11 or newer and uses ordinary scientific Python dependencies listed in \texttt{requirements.txt}. From the release root, the shortest public workflow is:

\begin{verbatim}
make reproduce-public
make test
make hygiene
\end{verbatim}

The first command regenerates the public WILD and source-registry outputs, the second runs the test suite, and the third verifies that the public release does not contain raw model responses, benchmark text, private parser-audit samples, or local cache paths. The package also includes a \texttt{manifest} target that refreshes file hashes for the public release.

\begin{table}[t]
\centering
\small
\begin{tabular}{ll}
\toprule
Field & Example output \\
\midrule
Current model & Qwen2.5-1.5B-Instruct \\
Candidate model & Qwen2.5-3B-Instruct \\
Shared WILD items & 54,464 \\
Accuracy delta & +5.7 percentage points \\
Improvement mass & 18.0\% \\
Regression mass & 12.3\% \\
Churn mass & 30.4\% \\
Interpretation & Near-parity aggregate gain with substantial item-level churn \\
\bottomrule
\end{tabular}
\caption{Worked WILD audit-card example for a selected-task pooled same-family comparison. The example illustrates how a small aggregate gain can still require reviewing a large changed item set.}
\label{tab:worked-audit-card}
\end{table}

\section{Use Case 1: WILD Correctness-Only Audit}

The WILD demonstration is the parser-independent empirical support for the audit protocol because it avoids answer extraction. WILD exposes item identifiers, model identifiers, task identifiers, and binary correctness records. The full WILD release contains evaluations of 65 models over 109,564 unique items spanning 163 tasks from 27 datasets \parencite{krumdick2026wild}. The present artifact uses a selected public subset containing five Qwen and Llama models up to approximately four billion parameters and ten tasks. It supports item-level churn accounting, but it does not support raw-response parsing claims, prompt-format claims, or manual answer-extraction validation.

The WILD use case validates the pairwise accounting layer of the artifact, not the raw-output scoring layer. The subset rule is fixed in the artifact: retain public item-level correctness rows for same-family small-model chains with at least two public models at or below approximately four billion parameters, shared coverage over the selected tasks, and a plausible size or generation replacement path. The selected tasks are MMLU-Pro, MMLU, ARC-Challenge, GSM8K, HellaSwag, PIQA, BoolQ, IFEval, BBH, and MATH. Models without a retained same-family comparator are excluded from pairwise analysis but listed in Table~\ref{tab:wild-inclusion}. This is a public cross-benchmark demonstration of the audit protocol, not a representative analysis of all WILD models or all WILD tasks. No estimate in this section is interpreted as the average churn rate of all WILD models.

\begin{table}[t]
\centering
\scriptsize
\resizebox{\linewidth}{!}{\input{tables/wild_model_inclusion_coverage.tex}}
\caption{WILD inclusion coverage for public models at approximately four billion parameters or below. The table reports selected-task rows available under the artifact extraction rule and the reason each model was retained or excluded from the paired-family demonstration. The complete machine-generated inclusion and exclusion files are included in the public release.}
\label{tab:wild-inclusion}
\end{table}

Rows in Table~\ref{tab:wild-inclusion} are retained model-task correctness records before pairwise intersection; pairwise $N$ in Table~\ref{tab:wild-pairs} is the number of shared item identifiers after aligning each directed model pair.

\begin{table}[t]
\centering
\small
\input{tables/wild_replication_summary.tex}
\caption{Public WILD correctness-only demonstration.}
\label{tab:wild-summary}
\end{table}

The generated WILD replication contains 44 directed same-family comparisons: four selected-task pooled comparisons, plus 40 task-level comparisons across ten tasks. Across the selected WILD tasks, the three Qwen selected-task pooled directed comparisons have 34.4\% mean churn. The Qwen2.5-0.5B to Qwen2.5-3B comparison improves aggregate correctness by 22.2 percentage points while changing 40.9\% of item outcomes and regressing 9.3\% of all shared items. The near-parity Qwen2.5-1.5B to Qwen2.5-3B comparison improves aggregate correctness by 5.7 percentage points while changing 30.4\% of item outcomes. This is the core operational result in a parser-free setting: aggregate gain and item-level replacement churn are different quantities.

\begin{table}[t]
\centering
\small
\resizebox{\linewidth}{!}{\input{tables/wild_all_selected_pairs.tex}}
\caption{WILD pooled replacement audits. The table reports selected-task pooled directed comparisons for retained Qwen and Llama model chains. Values are computed from binary correctness records only and do not evaluate raw-response parsing. $I$ is improvement mass, $R$ is regression mass, and NRB is normalized regression burden.}
\label{tab:wild-pairs}
\end{table}

The Llama 3.2 1B to 3B comparison gives the same qualitative lesson. It improves by 17.8 percentage points while changing 35.3\% of item outcomes and regressing 8.8\% of all shared items. The audit therefore identifies not only whether a candidate model wins on average, but how much of the correctness pattern is rewritten.

\begin{table}[t]
\centering
\small
\resizebox{\linewidth}{!}{\input{tables/wild_task_dispersion.tex}}
\caption{Task-level dispersion for WILD directed comparisons. The table summarizes the 40 task-level comparisons, separate from the four selected-task pooled comparisons in Table~\ref{tab:wild-pairs}, and reports shared-item denominators.}
\label{tab:wild-task-dispersion}
\end{table}

The task-level dispersion table shows why an all-task mean is not enough. Across all selected WILD task-level comparisons, median churn is 33.9\% with an interquartile range of 30.0--39.4\%. For Qwen alone, median task-level churn is 32.6\%, and the highest mean-churn task among the selected tasks is PIQA at 42.9\%. These values do not imply a universal churn level for all small models. They show that the accounting method can be applied to a public item-level correctness matrix and that large churn appears outside the MMLU-Pro archived-output case study.

\section{Use Case 2: MMLU-Pro Raw-Output Diagnostic}

The MMLU-Pro case study uses 14 archived model-output runs over 3,008 shared items. A source-manifest script reconstructs the evaluated rows from the TIGER-Lab/MMLU-Pro test parquet snapshot and verifies saved identifiers against source category and answer labels. The manifest stores normalized item hashes and source identifiers, not benchmark question text. The reconstructed source revision is \nolinkurl{b189ec765aa7ed75c8acfea42df31fdae71f97be}; the source parquet SHA256 is \nolinkurl{0e24a191921c2f453518a537a8b2117bd137e7714d4ef1565e9ba06c1ecb9ad8}; and the public manifest item-hash SHA256 is \nolinkurl{1f482b880fad0d8c97aa08f83d250ffd43e715b78ec2619987e0db4e9a98d9d3}.

\begin{table}[t]
\centering
\small
\input{tables/mmlu_pro_manifest_status.tex}
\caption{MMLU-Pro source-manifest reconstruction. The public artifact stores row hashes and source identifiers, not benchmark question text or answer-option text.}
\label{tab:mmlu-pro-manifest}
\end{table}

All 3,008 evaluated identifiers match the source rows with zero category or answer mismatches. The subset is now reconstructable, but the original reason for selecting exactly these 3,008 rows from MMLU-Pro was not preserved. Because the subset-selection rationale was not preserved, no sampling inference is made from this subset. The subset is used only to exercise the raw-output audit pipeline.

The raw-output diagnostic illustrates what is required beyond a binary correctness matrix. Three runs include saved prediction fields; the remaining rows require deterministic extraction from raw responses. A conservative prompt-echo guard marks responses ending in a bare chat marker after echoed prompt text as unanswered. This matters for Gemma-4-E2B-Instruct: after the guard, its answered rate is 5.2\% and parse-failure rate is 94.8\%. That run is retained in descriptive files as a generation-format failure but excluded from primary pairwise claims.

The parser-audit sample has been generated but not labeled. Therefore raw-output parser validation is not claimed, and the MMLU-Pro results are not used as confirmatory evidence for model-family behavior. Table~\ref{tab:mmlu-diagnostic-status} summarizes the diagnostic status of the case study.

\begin{table}[t]
\centering
\small
\begin{tabular}{p{0.32\linewidth}p{0.24\linewidth}p{0.34\linewidth}}
\toprule
Component & Status & Main-claim use \\
\midrule
Source reconstruction & 3,008/3,008 item identifiers matched & Provenance support \\
Raw-response extraction & Deterministic parser plus scoring-mode sensitivity & Diagnostic only \\
Manual parser validation & Audit sample generated; labels incomplete & Not claimed \\
MMLU-Pro family comparison & Parser-dependent and incompletely manifested & Appendix diagnostic only \\
\bottomrule
\end{tabular}
\caption{Diagnostic status of the archived MMLU-Pro raw-output workflow. The WILD-supported core claim does not depend on MMLU-Pro parser validation.}
\label{tab:mmlu-diagnostic-status}
\end{table}

The scoring-mode sensitivity check uses the three runs that contain both saved predictions and raw responses. Table~\ref{tab:scoring-mode} compares the default saved-prediction-primary scoring with a raw-response-only parser. It shows that scoring asymmetry is materially important, not merely measurable: Gemma-3-1B-Instruct changes correctness on 11.6\% of rows across the two modes, and Phi-2's raw-response-only answerability is much lower than its saved-prediction-primary answerability. This sensitivity check is not a substitute for manual validation, but it prevents the paper from treating the parser as invisible.

\begin{table}[t]
\centering
\small
\resizebox{\linewidth}{!}{\input{tables/scoring_mode_sensitivity.tex}}
\caption{Scoring-mode sensitivity for runs with both saved predictions and raw responses. Values are percentages. $\Delta$ is raw-response-only accuracy minus saved-prediction-primary accuracy.}
\label{tab:scoring-mode}
\end{table}

The archived raw run files preserve item IDs, categories, ground-truth labels, model labels, and raw responses. They do not preserve exact Hugging Face repository identifiers, revision hashes, prompt templates, chat templates, decoding parameters, runtime versions, run dates, or complete hardware metadata. Consequently, the MMLU-Pro accounting can be regenerated from saved outputs, but the original inference runs cannot be independently reproduced as validated model-behavior evaluations without rerunning or reconstructing the experiment.

Compact MMLU-Pro extraction coverage, parser gates, run-metadata status, scoring sensitivity, and one same-family diagnostic summary are reported in the appendix. Full directed comparison tables, domain triage outputs, near-parity diagnostics, and generated figures are included in the public release rather than promoted in the manuscript PDF. They document artifact behavior and do not support confirmatory MMLU-Pro model-evaluation claims.

\section{External Evidence Availability}

The artifact classifies external evidence by what it can prove. This distinction is not bookkeeping; it is the boundary between valid replacement accounting and unsupported extrapolation. Error persistence, regression mass, and churn require paired item-level outcomes. Aggregate leaderboards can contextualize model quality, but they cannot identify corrected errors or newly introduced regressions.

\begin{table}[t]
\centering
\small
\resizebox{\linewidth}{!}{\input{tables/external_claim_use_table.tex}}
\caption{External evidence map used by the artifact. Only Tier A/B item-level sources can support replacement-accounting metrics. Tier C sources are context only, and Tier D sources are logged as exclusions unless access is later granted.}
\label{tab:external-claim-use}
\end{table}

Public model cards and technical reports provide useful aggregate context for small-model capability \parencite{qwen25technicalreport,qwen3technicalreport,gemma2technicalreport,gemma3technicalreport,llama3herd}, but they are not substitutes for paired item-level outputs. The artifact records source-backed aggregate scores and source pointers as context-only evidence in \texttt{analysis/external\_benchmark\_context/}. These scores may differ in prompt format, shot count, model variant, benchmark revision, and evaluation harness. For that reason, no paper claim about error persistence, regression mass, or churn is derived from external aggregate scores.

\section{Discussion}

\subsection{What the evidence supports}

The public WILD demonstration supports the core audit claim without answer extraction: aggregate gains can coincide with large item-level churn. The MMLU-Pro diagnostic is included to show how the artifact handles raw-response evidence with incomplete scoring validation. Family-level diagnostic summaries are provided only in the appendix and public release to document artifact behavior; they are not part of the article's core claim set.

\subsection{What the evidence does not support}

The evidence does not support a universal scaling law, a mechanistic explanation of why families differ, or a claim that one family is generally better for deployment. It also does not support interface-tax, long-context, safety, tool-use, multimodal, or multi-turn claims. The WILD results are correctness-only. The MMLU-Pro results are parser-dependent until the manual audit is complete.

\subsection{How practitioners should use the audit}

For an actual deployment, the audit should be run on the proposed successor path and on the task distribution that matters for the product. A high-churn replacement may be acceptable when corrected items are more important than regressions. That is a product and risk decision, not a conclusion implied by aggregate accuracy.

The audit card should be treated as a triage tool. Large regression mass, high normalized regression burden, and near-parity churn should trigger manual inspection of affected item clusters before deployment.

\section{Threats to Validity}

\paragraph{Construct validity.}
The audit measures changes in binary correctness over shared labeled items. It does not directly measure usefulness, calibration, safety, open-ended quality, or user satisfaction. WILD avoids parser error but is limited to correctness records. MMLU-Pro uses raw outputs but depends on deterministic answer extraction.

\paragraph{Parser validity.}
The MMLU-Pro parser may misclassify some raw completions, especially fallback cases. The generated 452-row manual audit sample includes 120 high-risk rows, but it must still be human-labeled before the raw-output case study can be treated as a fully validated model-evaluation result. This is the primary blocker for treating the MMLU-Pro case study as validated model evaluation. Until then, MMLU-Pro churn and regression claims are correctness-under-parser claims. The saved-prediction sensitivity check shows that scoring mode can change per-row correctness for runs with both saved predictions and raw responses.

\paragraph{Internal validity.}
The archived MMLU-Pro run logs do not preserve exact repository identifiers, model revision hashes, prompt templates, chat templates, decoding parameters, runtime versions, run dates, or complete hardware metadata. The reported accounting can be regenerated from saved outputs, but model-behavior replication requires full run manifests or reruns.

\paragraph{External validity.}
The primary raw-output case study uses one fixed MMLU-Pro subset. WILD adds public cross-benchmark correctness evidence over a selected subset of models and tasks, but it does not include the same prompts, raw completions, or parser decisions. Additional raw-output datasets and model families are needed for stronger generalization.

\paragraph{Selection and direction.}
The MMLU-Pro all-pairs diagnostic uses a deterministic ordering rule, not a complete vendor release chronology. Directed comparisons marked as chronology unclear should not be interpreted as real deployment replacements.

\section{Artifact Outputs and Reproduction}

The artifact is organized around generated, inspectable files rather than hand-computed paper values. The MMLU-Pro pipeline emits pairwise metrics as CSV/JSON, replacement-audit cards as CSV/JSON/Markdown, quality-gate summaries, parser-audit summaries, scoring-mode sensitivity tables, source-manifest hashes, and paper-ready LaTeX tables. The WILD pipeline emits normalized correctness records, pairwise replacement metrics, task-family summaries, task-dispersion summaries, a source-coverage report, and paper-ready tables. The external-evidence pipeline emits a source-quality map, aggregate context tables, informal-context exclusions, and a claim-use manifest.

The artifact also includes a CPU-only mini-rerun harness for fully manifested follow-up validation. It writes a run manifest, environment file, selected-item hash manifest, resumable raw records, per-model summaries, pairwise audit metrics, and claim checks. A completed real-model mini-rerun is not used as evidence in the present manuscript unless its outputs, exact model revisions, prompt template, tokenizer/chat-template metadata, decoding settings, runtime environment, hardware summary, and run date are archived with the release.

The release also includes tests for source-manifest matching, parser-audit summaries, scoring-mode sensitivity, WILD replacement accounting, evidence-tier restrictions, CPU-rerun resume behavior, and public-release hygiene. A hygiene scan checks that public artifacts do not contain raw model responses, benchmark text, local cache paths, or private files.

\section{Data and Software Availability}

The GitHub/SSRN public proof-of-analysis package is available at \url{https://github.com/DsChauhan08/error_persistence_regression_gap}. The package contains the manuscript, analysis code, derived MMLU-Pro pairwise metrics, source-row hashes, parser-audit metadata, scoring-mode sensitivity summaries, WILD correctness-only replication outputs, evidence-tier manifests, generated tables, and generated figures. It does not redistribute full MMLU-Pro question text, full answer-option text, raw model responses, or item-level response excerpts.

\begin{table}[t]
\centering
\small
\resizebox{\linewidth}{!}{\input{tables/artifact_release_status.tex}}
\caption{Artifact availability and claim scope. Parser validation and MMLU-Pro confirmatory status remain visible but do not block the parser-independent WILD/source-manifest claim set for this software-tool article.}
\label{tab:artifact-release-status}
\end{table}

The accounting can be regenerated from saved outputs, but full model-behavior replication requires complete run manifests. A future rerun intended to support confirmatory model-evaluation claims should archive exact model repository IDs, revision hashes, prompt templates, chat templates, decoding parameters, inference engine, package versions, hardware summary, and run dates. The absence of completed manual parser labels is therefore not hidden or bypassed: parser-dependent MMLU-Pro claims are excluded from the core claim set in Table~\ref{tab:claim-ledger}.

The main public-package reproduction commands are:

\begin{verbatim}
make reproduce-public
make test
make hygiene
make manifest
\end{verbatim}

The \texttt{make reproduce-public} target regenerates the public WILD and evidence-registry outputs. The public package also exposes the underlying Python modules for users who want to rerun individual steps directly.

Parser-audit regeneration and scoring-mode sensitivity over private/source raw outputs require access to the withheld raw model-output JSONL files:

\begin{verbatim}
PYTHONPATH=src python -m boundary_slm.scoring_mode_sensitivity
PYTHONPATH=src python -m boundary_slm.parser_audit_impact
PYTHONPATH=src python -m boundary_slm.parser_audit_second_pass
PYTHONPATH=src python -m boundary_slm.mmlu_scoring_robustness
\end{verbatim}

The public package keeps only redacted audit metadata and aggregate summaries.

\section{Conclusion}

This paper presents an item-level audit for small language model replacement. Aggregate accuracy says whether a candidate model wins on average. The proposed decomposition shows how it wins, what it fails to fix, and what it breaks.

The public WILD demonstration shows that the audit can be applied to independent item-level correctness data and that modest aggregate gains can still involve large churn. The MMLU-Pro archived-output diagnostic illustrates how the same audit workflow applies to raw-response evidence, while exposing the need for source manifests, parser validation, scoring-mode sensitivity, and full run provenance.

The practical recommendation is simple: benchmark-based model replacement decisions should report corrected errors, regressions, persistent errors, and churn mass, not only aggregate accuracy.

\section{Author Contributions}

Dhananjay Singh Chauhan designed the study, ran or assembled the model-output evaluations, implemented the analysis code, performed the statistical analysis, generated the tables and figures, and wrote the manuscript.

\section{Competing Interests}

The author declares that there are no competing interests.

\section{Grant Information}

The author declares that no external funding supported this work.

\section{Acknowledgments}

The author thanks Kaggle and Google Colab for making accessible compute environments available for the MMLU-Pro testing and related validation runs. The author also thanks the maintainers of the open-source model, benchmark, and evaluation ecosystems that made the raw model-output and public correctness-matrix analyses possible.

\printbibliography

\appendix

\section{MMLU-Pro Diagnostic Details}

\begin{table}[H]
\centering
\small
\resizebox{\linewidth}{!}{\input{tables/extraction_coverage.tex}}
\caption{Answer-source coverage by family. Saved predictions are used when present; otherwise explicit answer markers and fallback letter rules are applied to raw responses.}
\label{tab:extraction-coverage}
\end{table}

\begin{table}[H]
\centering
\small
\input{tables/parser_validation_status.tex}
\caption{Parser-validation status for the archived MMLU-Pro raw-output diagnostic. Manual labels are required before reporting parser agreement as a validation result.}
\label{tab:parser-validation-status}
\end{table}

\begin{table}[H]
\centering
\small
\resizebox{\linewidth}{!}{\input{tables/parser_audit_impact.tex}}
\caption{Parser-audit impact gate. The current artifact has generated the sample but has not completed human labels, so MMLU-Pro remains a parser-dependent diagnostic.}
\label{tab:parser-impact}
\end{table}

\begin{table}[H]
\centering
\small
\resizebox{\linewidth}{!}{\input{tables/mmlu_scoring_robustness.tex}}
\caption{MMLU-Pro scoring-robustness modes over archived outputs. These are sensitivity analyses, not confirmatory validation while parser labels are incomplete.}
\label{tab:mmlu-scoring-robustness}
\end{table}

\section{MMLU-Pro Model Runs and Quality Gate}

\begin{table}[H]
\centering
\small
\resizebox{\linewidth}{!}{\input{tables/model_quality_gate.tex}}
\caption{Model-level leaderboard and quality-gate status for the archived MMLU-Pro outputs. Accuracy and answered values are percentages over the same 3,008-item set.}
\label{tab:model-quality-gate}
\end{table}

\section{MMLU-Pro Run Metadata Status}

\begin{table}[H]
\centering
\small
\resizebox{\linewidth}{!}{\input{tables/model_run_metadata_status.tex}}
\caption{Run-metadata completeness in the archived MMLU-Pro raw-output files.}
\label{tab:model-run-metadata}
\end{table}

\section{MMLU-Pro Same-Family Diagnostic Summaries}

\begin{table}[H]
\centering
\small
\resizebox{\linewidth}{!}{\input{tables/family_summary.tex}}
\caption{Primary MMLU-Pro same-family diagnostic accounting after excluding runs with answered rate below 80\%. Values are percentages. $I$ is improvement mass, $R$ is regression mass, and NRB is normalized regression burden.}
\label{tab:mmlu-family}
\end{table}

Full MMLU-Pro directed comparison tables, near-parity diagnostics, domain triage outputs, descriptive family checks, and generated figures are included in the public release under \texttt{analysis/paper\_metrics/} and \texttt{paper/figures/}. They are not part of the core WILD-supported claim set.

\section{Answer-Extraction Rules}
\label{app:parser-rules}

Answer extraction is deterministic. If a row contains a saved \texttt{prediction} field equal to one valid option letter, that field is used. Otherwise the parser searches the raw response in the following order and returns the last matching valid option letter:

\begin{enumerate}
  \item A single-letter response, optionally wrapped in parentheses or followed by a period.
  \item Explicit final-answer markers: \texttt{final answer}, \texttt{correct answer}, \texttt{correct option}, \texttt{correct choice}, \texttt{therefore}, \texttt{answer is}, and \texttt{option is}.
  \item Tail option markers in the final 500 characters of the response, such as a standalone option letter followed by punctuation.
  \item The final standalone option letter in the response, used only as a low-confidence fallback.
\end{enumerate}

Before these rules are applied, a prompt-echo guard marks a response as unanswered if it ends in a bare chat marker after echoed prompt text and contains no explicit answer marker. The valid option alphabet is \texttt{A}--\texttt{J}, matching the saved ground-truth labels. Rows with no extracted valid option are counted as unanswered and incorrect.

\end{document}
